---
title: "Assignment 1"
author: "Chhiring Lama"
date: "2024-01-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(tidyverse)
library(dplyr)
```

## Import data file as CSV

```{r}
my_data <- suppressWarnings(read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv", header = TRUE))
```

Disply Overview of Dataframe
 
```{r}
glimpse(my_data)
```
Extract a subset from the Data-frame with certain columns. 

```{r}
my_data1 <- subset(my_data, select = c(age, n, alcohol_use,marijuana_use))
as.data.frame(my_data1)
```
Extract only the row with age group of 50-54 that is baby boomers generation.

```{r}
boomers_50_64 <- my_data1 %>%
  filter(age == "50-64")
  
```

Filter all rows which are below 21 and assign them in boomers_50_64 variable.

```{r}
underAgeDrugUse <- my_data1 %>% 
    filter(age <= 21) 
underAgeDrugUse
```

Display baby boomers row.

```{r}
boomers_50_64

```
Compare two age groups those who use alcohol and marijuana.

```{r}
drug_comparison <- data.frame(
  age_group = c("Below 21", "50-64"),
  alcohol_use = c(mean(my_data1$alcohol_use[my_data1$age < 21], na.rm = TRUE), my_data1$alcohol_use[my_data1$age == "50-64"]),
  marijuana_use = c(mean(my_data1$marijuana_use[my_data1$age < 21], na.rm = TRUE), my_data1$marijuana_use[my_data1$age == "50-64"])
  
)
```

```{r}
drug_comparison
```
Bar Plot for alcohol use

```{r}
ggplot(drug_comparison, aes(x = age_group, y = alcohol_use, fill = age_group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Alcohol Use Between Age Groups",
       x = "Age Group",
       y = "Mean Alcohol Use") +
  theme_minimal()

```
Bar plot for marijuana use.

```{r}
ggplot(drug_comparison, aes(x = age_group, y = marijuana_use, fill = age_group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Marijuana Use Between Age Groups",
       x = "Age Group",
       y = "Mean Marijuana Use") +
  theme_minimal()
```

In summary, comparing alcohol and marijuana use between individuals under 21 and the baby boomer generation (50-64 age group) reveals that, on average, those under 21 report higher marijuana use, while baby boomers report higher alcohol use. These findings emphasize the need for age-specific insights in shaping interventions and policies. Considering the limitations of mean values, further analysis may require for a comprehensive understanding of drug use patterns across age groups. Additionally, exploring the correlation between baby boomers and alcohol use may be a subject for future study.
